---
title: "Monte Carlo reliability calculation of a parallel system with correlated times to failure using copulas"
author: "Michy Alice"
date: "07 February 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The idea for this work is the following

- Find a system whose components may be affected by correlation between the times to failure (**tfs** for brevity in the following).
- Determine correlation structure between the **tfs** of the different components by estimating the copula (or choose which copula to use a priori) from experimental data collected from the system.
- Estimate copula type and parameters (or choose them a priori).
- Perform Monte Carlo evaluation of system reliability.
- Compare the results with the case of independent **tfs**.
- Draw appropriate conclusions.

## Example of application to a simple parallel system

Only reliability and MTTF are computed even though the analysis may be targeted depending on the mission.

### System description
The system has only 2 components in parallel working in the same operating conditions.

This script aims to model the case where tf1 and tf2 are correlated (but not necessarily dependent).

The correlation structure (usually called in the literature, "dependence structure") is given by the copula.
 
The marginals (i.e. the distribution of the time to failure of each of the components) are independent of the copula. This means that, from a modelling perspective, the same correlation structure can be applied to components with different distribution of the time to failure.
Different correlation structures can be considered by using different copulas (for example: Gaussian, Gumbel, Frank, etc...). As a sidenote, the case of independence between the two failure rates can be obtained by using an independence copula.

We assume the following:

- The two components have a failure rate distributed according to an exponential distribution with parameter 0.01 years^(-1)
- The correlation structure is assumed to be a *t-copula* with parameters $df = 1$ and main parameter equal to $rho.1 = -0.8$.

The following code sets up the system and samples the time to failure of each component from the selected copula and marginals.

```{r}
#-------------------------------------------------------------------------------
# Boolean table of the system

# Parallel system
bool_table <- matrix(c(0, 0, 0, 0, 1, 1, 1, 0, 1, 1, 1, 1), nrow=4, ncol=3, byrow = T)

#-------------------------------------------------------------------------------

# Time horizon considered for reliability calculation [years]
T_HORIZON <- 450
# Time mission considered [years]
T_M <- 200
# Time for calculating conditional reliability [years]
T_COND <- 20
# Number of Monte Carlo iterations
N_MC_SIM <- 5000
# Number of system components
N_COMP <- ncol(bool_table) - 1
# For reproducibility
set.seed(2561)
# System state matrix
system_state_matrix <- matlab::zeros(N_MC_SIM, T_HORIZON + 1)
# Fix equal to 1 element k, 1 of the state matrix for each row since R(0)=1
system_state_matrix[1:N_MC_SIM, 1] <- 1
# All possible existing combinations of system state
ALL_POSSIBLE_STATES <- bool_table[, 1:N_COMP]
# Working and not working states: working statets are those with 1
SYS_STATES <- bool_table[, N_COMP + 1]

#-------------------------------------------------------------------------------
# Failure rate distribution modelling and correlation structure

# Failure rate of each component [years^(-1)]
LAMBDA = 0.01
# Correlation structure
COPULA_TYPE <- copula::tCopula(param = -0.8, df = 1, dim = N_COMP)
#copula::normalCopula(param=-0.9, dim = N_COMP)#copula::indepCopula(dim = N_COMP)

# Can try different margins (for instance weibull)
MARGINS <- rep("exp", N_COMP)
PARAM_MARGINS <- rep(list(rate=LAMBDA), 2)

# Multivariate distribution from copula
my_dist <- copula::mvdc(COPULA_TYPE,
                        margins = MARGINS,
                        paramMargins = PARAM_MARGINS)

# Sample N_MC_SIM time to failure from the distribution
sim_ttf <- copula::rMvdc(N_MC_SIM, my_dist)
```

## Correlation structure visualization

This is a representation of the sampled values of **tf1** and **tf2**. It can be noted that the failure rates are negatively correlated.

```{r}
psych::pairs.panels(sim_ttf, method="pearson", labels=c("tf1", "tf2"))
```


## Monte Carlo simulation of sistem reliability

Using this model, a Monte Carlo simulation is used to compute the reliability of the system and compare the results with the case of independent failure rates ( $R_s(t) = 2 R(t) - R(t)^2$ ).

```{r}
#-------------------------------------------------------------------------------
# Monte Carlo iterations

# Monte Carlo iterations outer loop
for(k in 1:N_MC_SIM)
{
    # Components state matrix: Ncomp x (T_horizon + 1)
    components_state_matrix = matlab::zeros(N_COMP, T_HORIZON + 1)
    # Fix equal to 1 element (k, 1) of the state matrix for each row k
    # since reliability is 1 at beginning of the period
    components_state_matrix[1:N_COMP, 1] = 1

    # Sample tf for each component
    for(i in 1:N_COMP)
    {
        # Sample tf (tf is in years)
        #tf = rexp(1, rate=LAMBDA)
        tf = sim_ttf[k, i]
        # From year 1 to ceil(tf) component has worked, set 1.
        # Note that index is shifted! Year 1 is at index 2.
        components_state_matrix[i, 2:min(ceiling(tf), T_HORIZON + 1)] = 1
        # From year ceil(tf) to year T_m component has not worked, set 0
        components_state_matrix[i, min(ceiling(tf) + 1, T_HORIZON + 1):(T_HORIZON + 1)] = 0
    }
    
    # Determine system history.
    for(i in 2:(T_HORIZON + 1))
    {
    
        # State system at time i is equal to column i of
        # components_state_matrix
        current_system_state = components_state_matrix[1:N_COMP, i]
        
        # Find if current_system_state is a working state
        for(j in 1:2^N_COMP)
        {
            # This finds the corresponding state from all possible states
            index_sum = sum(current_system_state == ALL_POSSIBLE_STATES[j, 1:N_COMP])
            # Once found the corresponding state we exit the loop
            if(index_sum == N_COMP)
            {
                break
            }
        }
    
        # Corresponding state is the j-th element of SYS_STATES
        system_status = SYS_STATES[j]
    
        # If system_status is 0 system is not working
        if(system_status == 0)
        {
            system_state_matrix[k, i:(T_HORIZON+1)] = 0 
            break
        }
    
        # Else the system is still working
        system_state_matrix[k, i] = 1
    }
}

#-------------------------------------------------------------------------------
# Calculate quantities of interest

# System reliability R(t)
R <- colSums(system_state_matrix)/N_MC_SIM

# MTTF: integrating with unit increment
MTTF <- sum(R)

# Print results

# Print MTTF
paste('MTTF correlated case:', MTTF, 'years')
paste('MTTF independent case:', 1.5*(1/LAMBDA), 'years')

# Estimated probability that system survives 20 years
paste('Estimated reliability at 20 years', R[21])

#-------------------------------------------------------------------------------
# Plot results

# Plot of reliability
plot(0:T_HORIZON, R, type="l", xlab = 'Time [years]', ylab = 'R(t)',
     main='System Reliability', xlim=c(0 , T_HORIZON), ylim=c(0, 1), col="red")

# Reliability in conditions of indepence betweeen the failure rates: analytical expression
lines(0:T_HORIZON, 2*exp(-LAMBDA*0:T_HORIZON) - exp(-2*LAMBDA*0:T_HORIZON), col="green")

legend("topright", legend=c("R(t) correlated case (negative correlation)",
                            "R(t) tfs independent"),
       col=c("red", "green"), lty=rep(1, 2), cex=0.8)
```

## Observations

Two observations can be done:

- Reliability increases in a specific time interval due to the negative correlation between the two **tfs**.
- MTTF increases by about 11 years with respect to the independent case (in this case!).

---

# Reliability and MTTF vs copula parameter

## System reliability vs copula parameter

By changing the correlation between the two *tfs*, reliability and MTTF both change. Changes in the correlation structure are reflected in changes of the copula parameters. In the following, a sweep of the t-copula parameter is performed. The parameter **rho.1** is swept from -0.9 to 0.9 by 0.1 increments.

```{r, echo=FALSE}

simulate_monte_carlo <- function(cop)
{
    #-------------------------------------------------------------------------------
    # Boolean table of the system
    
    # Parallel system
    bool_table <- matrix(c(0, 0, 0, 0, 1, 1, 1, 0, 1, 1, 1, 1), nrow=4, ncol=3, byrow = T)
    
    #-------------------------------------------------------------------------------
    
    # Time horizon considered for reliability calculation [years]
    T_HORIZON <- 450
    # Number of Monte Carlo iterations
    N_MC_SIM <- 5000
    # Number of system components
    N_COMP <- ncol(bool_table) - 1
    # System state matrix
    system_state_matrix <- matlab::zeros(N_MC_SIM, T_HORIZON + 1)
    # Fix equal to 1 element k, 1 of the state matrix for each row since R(0)=1
    system_state_matrix[1:N_MC_SIM, 1] <- 1
    # All possible existing combinations of system state
    ALL_POSSIBLE_STATES <- bool_table[, 1:N_COMP]
    # Working and not working states: working statets are those with 1
    SYS_STATES <- bool_table[, N_COMP + 1]
    
    #-------------------------------------------------------------------------------
    # Failure rate distribution modelling and correlation structure
    
    # Failure rate of each component [years^(-1)]
    LAMBDA = 0.01
    # Correlation structure
    COPULA_TYPE <- cop
    
    # Can try different margins (for instance weibull)
    MARGINS <- rep("exp", N_COMP)
    PARAM_MARGINS <- rep(list(rate=LAMBDA), 2)
    
    # Multivariate distribution from copula
    my_dist <- copula::mvdc(COPULA_TYPE,
                            margins = MARGINS,
                            paramMargins = PARAM_MARGINS)
    
    # Sample N_MC_SIM time to failure from the distribution
    sim_ttf <- copula::rMvdc(N_MC_SIM, my_dist)
    
    
    #-------------------------------------------------------------------------------
    # Monte Carlo iterations
    
    # Monte Carlo iterations outer loop
    for(k in 1:N_MC_SIM)
    {
        # Components state matrix: Ncomp x (T_horizon + 1)
        components_state_matrix = matlab::zeros(N_COMP, T_HORIZON + 1)
        # Fix equal to 1 element (k, 1) of the state matrix for each row k
        # since reliability is 1 at beginning of the period
        components_state_matrix[1:N_COMP, 1] = 1
        
        # Sample tf for each component
        for(i in 1:N_COMP)
        {
            # Sample tf (tf is in years)
            #tf = rexp(1, rate=LAMBDA)
            tf = sim_ttf[k, i]
            # From year 1 to ceil(tf) component has worked, set 1.
            # Note that index is shifted! Year 1 is at index 2.
            components_state_matrix[i, 2:min(ceiling(tf), T_HORIZON + 1)] = 1
            # From year ceil(tf) to year T_m component has not worked, set 0
            components_state_matrix[i, min(ceiling(tf) + 1, T_HORIZON + 1):(T_HORIZON + 1)] = 0
        }
        
        # Determine system history.
        for(i in 2:(T_HORIZON + 1))
        {
            
            # State system at time i is equal to column i of
            # components_state_matrix
            current_system_state = components_state_matrix[1:N_COMP, i]
            
            # Find if current_system_state is a working state
            for(j in 1:2^N_COMP)
            {
                # This finds the corresponding state from all possible states
                index_sum = sum(current_system_state == ALL_POSSIBLE_STATES[j, 1:N_COMP])
                # Once found the corresponding state we exit the loop
                if(index_sum == N_COMP)
                {
                    break
                }
            }
            
            # Corresponding state is the j-th element of SYS_STATES
            system_status = SYS_STATES[j]
            
            # If system_status is 0 system is not working
            if(system_status == 0)
            {
                system_state_matrix[k, i:(T_HORIZON+1)] = 0 
                break
            }
            
            # Else the system is still working
            system_state_matrix[k, i] = 1
        }
    }
    
    #-------------------------------------------------------------------------------
    # Calculate quantities of interest
    
    # System reliability R(t)
    R <- colSums(system_state_matrix)/N_MC_SIM
    
    # MTTF: integrating with unit increment
    MTTF <- sum(R)
    
    # Print results
    
    # Print MTTF
    paste('MTTF correlated case:', MTTF, 'years')
    paste('MTTF independent case:', 1.5*(1/LAMBDA), 'years')
    
    # Estimated probability that system survives 20 years
    paste('Estimated reliability at 20 years', R[21])
    
    return(R)
}

```

```{r, echo=FALSE}

# For reproducibility
set.seed(2561)

#-------------------------------------------------------------------------------
# T copula sweep

copulaparam <- seq(-0.9, 0.9, by = 0.1)
MTTFS <- numeric(length(copulaparam))
k <- 1
for(i in copulaparam)
{
    R <- simulate_monte_carlo(copula::tCopula(param = i, df = 1, dim = 2))
    if(i < -0.8)
    {
        plot(0:(length(R)-1), R, type="l", xlab = 'Time [years]', ylab = 'R(t)',
             main='System Reliability t-copula as rho1 varies', xlim=c(0 , 420), ylim=c(0, 1), col="red")
    }else
    {
        lines(0:(length(R)-1), R, col="red")
    }
    
    MTTFS[k] <- sum(R)
    k <- k + 1
}

legend("topright", legend=c("Correlated case(s)",
                            "Independent case"),
       col=c("red", "green"), lty=rep(1, 2), cex=0.8, lwd=c(1,2))

T_HORIZON <- 450
LAMBDA <- 0.01
lines(0:T_HORIZON, 2*exp(-LAMBDA*0:T_HORIZON) - exp(-2*LAMBDA*0:T_HORIZON), col="green", lwd=2)

plot(copulaparam, MTTFS, type="l", col="blue", lwd=2, main="MTTF vs t-copula parameter rho1", xlab="rho1")
lines(copulaparam, rep(150, length(copulaparam)), col="green", lwd=2)
legend("bottomleft", legend=c("MTTF correlated case",
                              "MMTF tfs independent (reference)"),
       col=c("blue", "green"), lty=rep(1, 2), cex=0.8, lwd=2)


```



```{r, echo=FALSE, eval=FALSE}
#-------------------------------------------------------------------------------
# Normal copula sweep

copulaparam <- seq(-0.9, 0.9, by = 0.1)
MTTFS <- numeric(length(copulaparam))
k <- 1
for(i in copulaparam)
{
    R <- simulate_monte_carlo(copula::normalCopula(param=i, dim = 2))
    if(i < -0.8)
    {
        plot(0:(length(R)-1), R, type="l", xlab = 'Time [years]', ylab = 'R(t)',
             main='System Reliability gauss copula as rho varies', xlim=c(0 , 420), ylim=c(0, 1), col="red")
    }else
    {
        lines(0:(length(R)-1), R, col="red")
    }
    
    MTTFS[k] <- sum(R)
    k <- k + 1
}

legend("topright", legend=c("Correlated case(s)",
                            "Independent case (reference)"),
       col=c("red", "green"), lty=rep(1, 2), cex=0.8, lwd=c(1,2))

T_HORIZON <- 450
LAMBDA <- 0.01
lines(0:T_HORIZON, 2*exp(-LAMBDA*0:T_HORIZON) - exp(-2*LAMBDA*0:T_HORIZON), col="green", lwd=2)

plot(copulaparam, MTTFS, type="l", col="blue", lwd=2, main="MTTF vs gaussian copula parameter rho", xlab="rho")
lines(copulaparam, rep(150, length(copulaparam)), col="green", lwd=2)
legend("bottomleft", legend=c("MTTF correlated case",
                            "MMTF tfs independent"),
       col=c("blue", "green"), lty=rep(1, 2), cex=0.8, lwd=2)


# #-------------------------------------------------------------------------------
# # What do samples look like?
# 
# for(i in seq(-0.9, 0.9, 0.1))
# {
#     COPULA_TYPE <- copula::normalCopula(param=i, dim = 2)
#     
#     # Can try different margins (for instance weibull)
#     MARGINS <- rep("exp", 2)
#     PARAM_MARGINS <- rep(list(rate=LAMBDA), 2)
#     
#     # Multivariate distribution from copula
#     my_dist <- copula::mvdc(COPULA_TYPE,
#                             margins = MARGINS,
#                             paramMargins = PARAM_MARGINS)
#     
#     # Sample N_MC_SIM time to failure from the distribution
#     sim_ttf <- copula::rMvdc(1000, my_dist)
#     
#     psych::pairs.panels(sim_ttf, method="pearson", labels=c("tf1", "tf2"))
# }

```

A similar sweep could be performed by keeping **rho.1** fixed and varying the parameter **df**.

---

## Next steps

The next steps can be the following

- Apply the model to a real system.
- ...
